---
title: "growing functions"
author: "Martin Monkman"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

# PUZZLE DU JOUR OF THE DAY

We have an incomplete series with values (colB), and the growth rate for the last value (colC).
Below is the tibble with 

```{r}
# test tibble; 10% compound annual growth rate
table_of_data <- tribble(
  ~colA, ~colB, ~colC,
  "a", 100, NA,
  "b", 110, NA,
  "c", NA, 0.1
)

```

```{r}
# this works...
table_of_data %>% 
  mutate(colB = case_when(
    is.na(colB) ~ lag(colB) * (1 + colC),
    TRUE ~ colB
  ))
```

--- 

Your challenge--make it work with more than one value. 
That is, the 121 in the third line above 
becomes the input that is now grown 


```{r}
# test tibble; 10% compound annual growth rate
table_of_data <- tribble(
  ~colA, ~colB, ~colC,
  "a", 100, NA,
  "b", 110, NA,
  "c", NA, 0.1,
  "d", NA, 0.2
)

# this does not work...
table_of_data %>% 
  mutate(colB = case_when(
    is.na(colB) ~ lag(colB) * (1 + colC),
    TRUE ~ colB
  ))

# tidyr::fill() will repeat the value in row 2 but...no arithmetic allowed
table_of_data %>% 
  fill(colB)
```

## loop!

```{r}

# I also tried "cumprod" but couldn't figure out the appropriate syntax for the partial series

### a FOR loop!

# iteration 1
output <- vector("double", nrow(table_of_data))
for (i in 1:nrow(table_of_data)) {
  output[i] <- table_of_data$colB[[i]]
}

# iteration 2
output <- vector("double", nrow(table_of_data))
for (i in 1:nrow(table_of_data)) {
  output[i] <- ifelse(is.na(table_of_data$colB[[i]]),
                      output[[i - 1]],
                      table_of_data$colB[[i]])
}

```

### third time lucky

inside a dataframe

```{r}

df_grow <- table_of_data
grow_var <- table_of_data$colB
# if growth is in column of variables
pct_var <- table_of_data$colC
# if growth is a single value to be repeated
pct_var <- rep(0.027, times = nrow(df_grow))

output <- vector("double", nrow(df_grow))
for (i in 1:length(output)) {
  output[i] <- ifelse(is.na(grow_var[[i]]),
                      output[[i - 1]] * (1 + pct_var[[i]]),
                      grow_var[[i]])
}

table_of_data_augmented <- table_of_data %>% 
  mutate(new_val = output)
table_of_data_augmented

```


### Stephanie's solution

```{r}
table_of_data %>%
  fill(colB) %>%
  mutate(colC = ifelse(is.na(colC), 1, 1+colC),
                colD = cumprod(colC),
                new_val = colB*colD)


```



## loop within group_by 

The challenge: grow "val_var" by consistent percentage, where
each is multipled by 1.1 within the group

i.e. ax 10 becomes ay 11 and then az 12.1

### table_of_data_2

```{r table_of_data_2}
# test tibble; 10% compound annual growth rate
table_of_data_2a <- tribble(
  ~alpha, ~omega, ~val_var,
#  "a", "x", 10,
  "a", "y", 20,
#  "b", "x", NA,
  "b", "y", NA,
#  "c", "x", NA,
  "c", "y", NA
)

# test tibble; 10% compound annual growth rate
table_of_data_2b <- tribble(
  ~alpha, ~omega, ~val_var,
  "a", "x", 100,
  "a", "y", 200,
  "b", "x", NA,
  "b", "y", NA,
  "c", "x", NA,
  "c", "y", NA
)

table_of_data_2b
```



```{r}


df_grow <- table_of_data_2a
df_grow

grow_var <- df_grow$val_var
### if growth is in column of variables
##pct_var <- table_of_data$colC
# if growth is a single value to be repeated
pct_var <- rep(0.1, times = nrow(df_grow))
pct_var

output <- vector("double", nrow(df_grow))
for (i in 1:length(output)) {
  output[i] <- ifelse(is.na(grow_var[[i]]),
                      output[[i - 1]] * (1 + pct_var[[i]]),
                      grow_var[[i]])
}

output

table_of_data_augmented <- df_grow %>% 
  mutate(new_val = output)
table_of_data_augmented


```


### 2b or not 2b

```{r}

this_one <- "x"

### if growth is in column of variables
##pct_var <- table_of_data$colC
# if growth is a single value to be repeated

df_grow <- table_of_data_2b %>% 
  filter(omega == this_one)

df_grow

grow_var <- df_grow$val_var


pct_var <- rep(0.1, times = nrow(df_grow))

output <- vector("double", nrow(df_grow))
for (i in 1:length(output)) {
  output[i] <- ifelse(is.na(grow_var[[i]]),
                      output[[i - 1]] * (1 + pct_var[[i]]),
                      grow_var[[i]])
}

output

table_of_data_augmented <- df_grow %>% 
  mutate(new_val = output)
table_of_data_augmented

    
table_of_data_2b_x <- table_of_data_2b %>% 
  left_join(table_of_data_augmented, by = c("alpha", "omega", "val_var"))

table_of_data_2b_x 

```


```{r}

this_one <- "y"

### if growth is in column of variables
##pct_var <- table_of_data$colC
# if growth is a single value to be repeated

df_grow <- table_of_data_2b %>% 
  filter(omega == this_one)

df_grow

grow_var <- df_grow$val_var


pct_var <- rep(0.1, times = nrow(df_grow))

output <- vector("double", nrow(df_grow))
for (i in 1:length(output)) {
  output[i] <- ifelse(is.na(grow_var[[i]]),
                      output[[i - 1]] * (1 + pct_var[[i]]),
                      grow_var[[i]])
}

output

table_of_data_augmented <- df_grow %>% 
  mutate(new_val = output)
table_of_data_augmented

    
table_of_data_2b_y <- table_of_data_2b_x %>% 
  left_join(table_of_data_augmented, by = c("alpha", "omega", "val_var"))

table_of_data_2b_y 

table_of_data_2b_y %>% 
  mutate(new_val = coalesce(new_val.x, new_val.y))


```

### that is the question

```{r}

fv(0.1, 3, 10)
clarklytics::create_par_value_tbl(10, 0.10, 0:2)

table_of_data_2b %>% 
  group_by(omega) %>% 
  mutate(val_var_2 = fv(0.1, 3, min(val_var)))

```


### Trying Stephanie's solution inside a group_by()

```{r}
table_of_data %>%
  fill(colB) %>%
  mutate(colC = ifelse(is.na(colC), 1, 1+colC),
                colD = cumprod(colC),
                new_val = colB*colD)

```


```{r}
growth_pct <- 0.1

table_of_data_2a %>%
  mutate(colC = ifelse(is.na(val_var), 1 + growth_pct, 1)) %>% 
  fill(val_var) %>%
  mutate(
    colD = cumprod(colC),
    new_val = val_var * colD)

```

And grouping...

```{r}

# applying Stephanie's solution to growing a series

# test tibble; 10% compound annual growth rate
df_data <- tribble(
  ~alpha, ~omega, ~val_var,
  "a", "x", 100,
  "a", "y", 200,
  "b", "x", NA,
  "b", "y", NA,
  "c", "x", NA,
  "c", "y", NA,
  "d", "x", NA,
  "d", "y", NA
)

growth_pct <- 0.1

df_data %>%
  group_by(omega) %>%
  # fill growth percentage column if the value variable is NA
  mutate(colC = ifelse(is.na(val_var), 1 + growth_pct, 1)) %>%
  # now fill value variable with constant value
  fill(val_var) %>%
  # calculate the cumulative growth rate (colD) at each step, then multiply that by the original value
  mutate(colD = cumprod(colC),
         new_val = val_var * colD)


```





maybe create 12 versions, one for each month

* filter for month, select demand and year_month

* run function

* then join back to main table on year_month
