---
title: "lottery odds"
author: "Martin Monkman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## lottery games 

Major lotteries have games that have the jackpot based on matching the numbers drawn. For example, one national lottery in Canada is **6-49** ... there are 49 balls numbered from 1 to 49, and 6 are randomly selected. If the numbers selected match the numbers played by an individual, that individual wins the jackpot. (There are smaller prizes based on matching 5, 4, and 3 numbers but that is outside the scope of this example.)


## 6-49

```{r}

ball_count <- 49
ball_select <- 6

ball_count_sequence <- seq(from = (ball_count - ball_select + 1), to = ball_count, by = 1)

permutations <- prod(ball_count_sequence) / factorial(ball_select)
permutations

```

Create a function

```{r}

lotto_perms <- function(ball_count, ball_select) {
  ball_count_sequence <- seq(from = (ball_count - ball_select + 1), to = ball_count, by = 1)
  permutations <- prod(ball_count_sequence) / factorial(ball_select)
  permutations
}

lotto_perms(49, 6)

```



## LottoMax

### original

```{r}

ball_count <- 50
ball_select <- 7

ball_count_sequence <- seq(from = (ball_count - ball_select + 1), to = ball_count, by = 1)

permutations <- prod(ball_count_sequence) / factorial(ball_select)
permutations

# 3 plays for $5
permutations / 3

```


```{r}
lotto_perms(50, 7)
```


### original

```{r}

ball_count <- 52
ball_game <- 7

ball_count_sequence <- seq(from = (ball_count - ball_select + 1), to = ball_count, by = 1)

permutations <- prod(ball_count_sequence) / factorial(ball_select)
permutations

# 4 plays for $6
permutations / 4


```

```{r}

lotto_perms(52, 7)

```

