---
title: "sorting columns"
author: "Martin Monkman"
date: "2023-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# need dplyr and other packages
library(tidyverse)

library(tidyselect)

```

## build sample data

Our first data set has a fairly straightforward structure: letter-number.

Note "c" before "b", "3" before "2"

```{r}

df1 <- tribble(
  ~a_1, ~c_1, ~b_1, ~x_1, ~a_3, ~c_3, ~b_3, ~a_2, ~c_2, ~b_2,
  1, 2, 3, 4, 5, 6, 7, 8, 9, 10
)

df1

```


## sort alphabetically

The `tidyselect::peek_vars()` pulls the names of the variables

* https://tidyselect.r-lib.org/reference/index.html


The `sort()` function sorts the names alphabetically

The `relocate`()` applies that 

* https://dplyr.tidyverse.org/reference/relocate.html


Note that this sorts based on the entire string--the "2" are now ahead of the "3"


```{r}

df1 |> 
  relocate(sort(peek_vars()))

```


To put "x1" first, specify that in the `select` function before the `sort()`


```{r}
df1_alpha <- df1 |> 
  select(x_1, sort(peek_vars()))

df1_alpha

```


### Sort in a custom order


Maintains the "c" before "b" / 3" before "2" ordering


```{r}
df1 |> 
  relocate(
    # put x_1 first
    x_1, 
    # then those that contain the pattern
    contains(match = "a_"), 
    # then the remainder in reverse alphabetical order
    sort(peek_vars(), decreasing = TRUE)
    )

df1_alpha |> 
  relocate(
    # put x_1 first
    x_1, 
    # then those that contain the pattern
    contains(match = "_3"), 
    # then the remainder in reverse alphabetical order
    sort(peek_vars(), decreasing = TRUE)
    )


```






Now...how to sort it numerically and alphabetically? (i.e. the original order)?







SEE ALSO

https://stackoverflow.com/questions/70008665/organize-columns-with-numerical-info-in-colnames-via-dplyr-relocate

https://codereview.stackexchange.com/questions/115217/order-columns-in-r-by-substring

```{r}
# extract variable names
df1_names <- names(df1)

str_split(df1_names, "_")

x1_values <- c("a", "b", "c", "x")
x2_values <- c(1:3)

df1_alpha_resort <- df1_alpha[order(factor(df1_alpha))]
  
  
```





Not going to work (I think)

https://stringr.tidyverse.org/reference/str_order.html


