---
title: "time zone check"
author: "Martin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)


## create a csv file with the full list of Olson Names
#ON_list <- as_tibble(OlsonNames())
#write_csv(ON_list, "Olson_Names.csv")

```

This code uses the {lubridate} function [`with_tz()`](https://lubridate.tidyverse.org/reference/with_tz.html) to convert a specific time to a different time zone.


## What time is it now?


```{r}

# local time
now()

# Tokyo
now(tz = "Asia/Tokyo")

```


```{r}
# assign now() to an object

current_time <- now()
current_time


# convert time
with_tz(current_time, tzone = "Asia/Tokyo")

```


## New York vs Tokyo vs Victoria

In this example, we first set a specific time object `time_to_convert`. This uses the {lubridate} function `ymd_hm()` to create a date-time object that is one minute past midnight in New York.

Next, the code uses the `with_tz()` function to change the time to Tokyo and Victoria, BC.


```{r}
# one minute past midnight
time_to_convert <- ymd_hm("2026-09-17 00:01", tz = "America/New_York")

print("Orlando time")
time_to_convert

print("Tokyo time")
with_tz(time_to_convert, tzone = "Asia/Tokyo")

print("Victoria time")
with_tz(time_to_convert, tzone = "Canada/Pacific")

```

