---
title: "time difference"
author: "Martin Monkman"
date: "2023-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
```


Winter solstice 2023 falls on December 21 in western Canada at 7:27 pm (Pacific Standard Time). This can also be written as "2023-12-22 03:27 UTC", using the Universal Time Coordinated standard. All times around the world are expressed relative to this standard.

* Time zones to the east are ahead, e.g. Japan is UTC +9 hours and New Zealand is in the last time zone before the International Date Line at UTC +12

* Time zones to the west are behind, e.g. Hawaii is UTC -10 and Pacific Standard Time is UTC -8

For this first section, we will use functions from the {lubridate} package.

```{r}
# lubridate functions
winter_solstice_utc <- lubridate::ymd_hm("2023-12-22 03:27", tz = "UTC")
winter_solstice_pst <- ymd_hm("2023-12-21 19:27", tz = "Canada/Pacific")

winter_solstice_utc - winter_solstice_pst 

#OlsonNames(tzdir = NULL)
```

## How long until the winter solstice occurs?

Now we have a precise time (to the nearest minute) of when the solstice will occur. How long is that from now?

Using lubridate function `now()`:

```{r}
# `now()` returns your computer's current time
now()

# subtract the winter solstice time from whatever now is"
winter_solstice_utc -  now()
```


```{r}
# NOTE: putting "now" first leads to negative value
# -- current date is a smaller value
now() - winter_solstice_utc

```

We can extract the value from "difference" using square brackets"

```{r}

ws_countdown <- winter_solstice_utc -  now()

ws_countdown_value <- ws_countdown[[1]]
ws_countdown_value

```

## Convert to days, hours, minutes

### days

```{r}
# `floor()` is a round-down to the nearest integer function (the opposite is `ceiling()`)
ws_countdown_days <- floor(ws_countdown_value)
ws_countdown_days

```

### hours and minutes

For hours, we have to subtract the days and then multiply by 24


```{r}
# calculate the remainder (the part after the decimal) after the whole number of days are subtracted
ws_countdown_value - ws_countdown_days

# multiply the remainder by 24 to get the number of hours
ws_hours_all <- (ws_countdown_value - ws_countdown_days) * 24
ws_hours_all

```

```{r}
ws_countdown_hours <- floor(ws_hours_all)
```


And now subtract the whole hours from the result and multiply by 60 to get the number of minutes

```{r}
# multiply that by 60 to get the number of minutes
ws_minutes_all <- (ws_hours_all - ws_countdown_hours) * 60
ws_minutes_all

ws_countdown_minutes <- floor(ws_minutes_all)
ws_countdown_minutes
```

And if we want even more precision, seconds


```{r}
# multiply that by 60 to get the number of hours
ws_countdown_seconds <- (ws_minutes_all - ws_countdown_minutes) * 60
ws_countdown_seconds

ws_countdown_seconds <- round(ws_countdown_seconds, 0)
ws_countdown_seconds
```

## put all the bits together into a single streamlined flow

```{r}
# calculate time difference, get value
ws_countdown_value <- (winter_solstice_utc -  now())[[1]]
# days
ws_countdown_days <- floor(ws_countdown_value)
# hours
ws_hours_all <- (ws_countdown_value - ws_countdown_days) * 24
ws_countdown_hours <- floor(ws_hours_all)
# minutes
# - note use `ceiling()` to round up partial minute
ws_minutes_all <- (ws_hours_all - ws_countdown_hours) * 60
ws_countdown_minutes <- ceiling(ws_minutes_all)


ws_countdown_days
ws_countdown_hours
ws_countdown_minutes
```

The winter solstice 2023 will occur in `r ws_countdown_days` days, `r ws_countdown_hours` hours, and `r ws_countdown_minutes` minutes.


