---
title: "duplication estimate"
author: "Martin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

If you have 23 people in the room, there is a 50/50 chance that 2 of them share a birthday.

Simulation:

* repeat a random selection of two items from a list (with replacement)

* calculate the proportion that are duplicated

## Toy version

Test with 6 sided dice

```{r}
# define the possible outcomes
die <- 1:6

# a random pair of numbers from the `die` object
dice <- sample(die, size = 2, replace = TRUE)

# returns row number of first duplicate
anyDuplicated(dice)

```

```{r}
# function
roll <- function() {
  die <- 1:6
  dice <- sample(die, size = 2, replace = TRUE)
  anyDuplicated(dice)

}

```

```{r}
# replicate 100 times
#set.seed(1729)
reps <- 100

rep_dupe_die <- 
  tibble(x = replicate(reps, roll()))

rep_dupe_die %>%
  filter(x > 0)

# calculate the proportion that are duplicated
rep_dupe_die %>% 
  filter(x > 0) %>% 
  nrow() / reps

```


## Birthdays

```{r}
# what proportion of a year is a single day?
1/365

# define the possible outcomes
days_in_year <- 1:365

# a random pair of numbers from the `days_in_year` object
birthday_table <- sample(days_in_year, size = 2, replace = TRUE)

# returns row number of first duplicate
anyDuplicated(birthday_table)

```

```{r}
# function
birthday_match <- function(num_dates = 2) {
  days_in_year <- 1:365
  
  # a random pair of numbers from the `days_in_year` object
  birthday_table <- sample(days_in_year, size = num_dates, replace = TRUE)
  
  # returns row number of first duplicate
  anyDuplicated(birthday_table)
}

```

```{r}
# replicate 100 times
#set.seed(1729)
reps <- 100000
number_of_people <- 23

rep_birthday <- 
  tibble(x = replicate(reps, birthday_match(number_of_people)))

rep_birthday %>%
  filter(x > 0)

# calculate the proportion that are duplicated
rep_birthday %>% 
  filter(x > 0) %>% 
  nrow() / reps


```


### With factorials

```{r factorial_365, eval=FALSE}

factorial(365)

```

Vnr = n! / (n - k)!
Vt = n^k = 365^23

n! / (n - k)! = factorial of seq(n to k+1)



```{r}

n = 10
k = 5

n_seq <- 1:n
k_seq <- 1:(n-k)

# concatenate into single vector
nk_seq <- c(n_seq, k_seq)

```

Reduce

* eliminate values that are common in both n_seq and k_seq

* https://stackoverflow.com/questions/37381174/r-removing-duplicate-elements-in-a-vector

* combine into single vector, remove all values that have a duplicate

```{r fun_keep_singles}
keep_singles <- function(v){
v[!(v %in% v[duplicated(v)])]
}


keep_singles(nk_seq)


```

Now with 365 and 23

```{r}

n = 365
k = 23

n_seq <- 1:n
k_seq <- 1:(n-k)

# concatenate into single vector
nk_seq <- c(n_seq, k_seq)


nk <- keep_singles(nk_seq)
nk


```

multiply the remaining values

```{r}
Vnr <- prod(nk)
Vnr
```

```{r}
Vt <- n^k
Vt

```

```{r}
1 - (Vnr / Vt)
```



## Fibonaci sequence

Is this a Fibonaci? 

* Value starts at zero, adds the value of the current row (eg 1st value is zero, then 2nd is 0 + 1 = 1, then 1 + 2 = 3, 3 + 3 = 6, etc)

```{r}
365/2

# number of cases
num_instance <- 23

output <- vector(mode = "integer", length = num_instance)  # 1. output

# the loop
j <- 0
# - "i" is the counter that 
for (i in 1:num_instance) {            # 2. sequence
  output[[i]] <- j   
  j <- (i + j)
}

# print the output object
output

```

